---
title: "Penguins ANOVA"
author: "Ilayda Dinc"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load packages and read -in penguins.csv data

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question: 

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring

New functions : with + table for easy counting over multiple factors

Update species name using case_when

```{r}
counts <- with(penguins,table(species,sex))

penguins2 <- penguins %>% 
  mutate(common_name =
           case_when(
             species=="Chinstrap penguin (Pygoscelis antarctica)" ~ "Chainstrap",
             species=="Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species=="Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% 
  filter(sex=="FEMALE")

penguins2


```

-Exploratory graphs
-Summary statistics
-Levene's test for equal variances
-one-way ANOVA
-post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x=flipper_length))+
  geom_histogram(aes(fill= common_name))+
  facet_wrap(~common_name)

qqs <- ggplot(penguins2, aes(sample= flipper_length))+
  geom_qq()+
  facet_wrap(~common_name)

qqs


# Levene's test for equal variances (>2 groups... if you have two groups, use the F-test for equal variances)

# We'll use the function leveneTest from the 'car' package

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)

# we retain the null hypothesis of equal variances (p= 0.06)

table <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance= var(flipper_length)
  )


```

#3. One-way ANOVA

one factor is species. there are 3 levels. random variable is flipper length

Single factor: common_name (species)
Number of Levels in the factor : 3 (Chainstrap, Gentoo, Adelie)
Random Variable: flipper_length (mm)

Is there a significant difference in mean flipper length(mm) for female Chainstrap, Gentoo and Adelie penguins?

```{r}
#H0: Mean flipper lengths across all species are equal
#HA: There is at least two means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data= penguins2)

summary(penguin_aov)


```
If we take samples from these three groups and they actually come from same populations, there is zero chance that 



At least two samples were taken from populations with different means.Which ones are different from each other? or something else?
```{r}
#post-hoc testing using Tukey's HSD
penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph
```

Mean flipper lengths for Adelie, Gentoo and Chinstrap penguins are all significantly different (F(df) = statistic,p<0.001, $\alpha$= 0.05 with post-hoc Tukey's HSD, $\alpha=0.05$)




